\begin{abstract}
    In the world of financial engineering, the use of smart contracts - programs which run on blockchain platforms and deal with cryptocurrencies - is becoming widespread. The implementation of financial contracts in the form of smart contracts is useful for automating the payments they obligate, and for keeping track of financial dealings for auditing purposes. Unfortunately, due to the high complexity of existing smart contract languages, these smart contracts are difficult to implement and evaluate, and can often contain vulnerabilities. \\
    
    Long before blockchains became commonplace, a functional domain-specific language for the declaration of financial contracts was created by Simon Peyton Jones, Jean-Marc Eber, and Julian Seward\cite{SPJ}. This DSL was devised to make financial contracts less verbose and complex, easier to evaluate mathematically, and less error-prone. \\
    
    We present a smart contract implementation of a slightly modified version of the aforementioned DSL, called \textit{SmartFin}. A single smart contract is implemented that can represent any SmartFin financial contract passed into its constructor. We also present a web client for composing and evaluating SmartFin financial contracts, and deploying and interacting with their corresponding smart contract instances. As such, a user need only compose a SmartFin financial contract, and is not required to program a smart contract. This allows smart contract representations of financial contracts to be created and evaluated with ease, and the simple syntax and semantics of SmartFin make implementation errors significantly less common in comparison to writing a bespoke smart contract.
\end{abstract}